\Title

### 2020-02-17 manuscript/manuscript.tex
Improve source distributions text

How to choose how many sources to use and where to locate them are not
trivial decisions which play an important role on the accuracy of the
predictions and the computational resources needed to fit the sources'
coefficients.

@@

A high number of sources scattered on the survey region is capable of
reproducing the observed data. Nevertheless, too many sources will make the
Jacobian matrix too large, what could introduce computational problems like
the inability of storing the matrix on memory or a high computational
time needed to fit the sources' coefficients.
On the other hand, using very few sources will undoubtedly reduce the memory
needs and lower the computational time when fitting the sources' coefficients.
Nevertheless, it could be incapable of making accurate predictions or even
reproducing the measured data if the number of sources is too low.

The choice of the source distribution could also depend on the characteristic
of the survey and how the observation points are distributed on the region.

A ground survey usually consists in observation points located along irregular
paths plus some possible scattered points. The survey region is usually
heterogeneously covered, leaving areas without any observation.

On the other hand, an airborne survey consists in observation points located
along almost straight flight lines. The measurements are usually taken at
a high frequency, so the observation points along the flight lines are very
close to each other. This creates an anisotropy on the observation points:
there are much more observation points along the flight lines directions.
Airborne surveys homogeneously cover the survey region.

@@

The \emph{regular grid} creates an homogeneous distribution of point sources
bellow the surveyed region.
It could create too many sources on areas without data and very few ones on
areas with a high number of observation points, leading to an underfitting of
the observed data.
This could be solved by reducing the grid spacing and therefore increasing the
number of point sources and the computational load of the interpolation
process.

The \emph{source beneath data} layout adds sources where they are more needed.
If the survey is composed by a large number of data points clustered on paths
(like an airborne survey), putting one source beneath each observation point
will create an anisotropic source distribution: there will be several sources
distributed along a privileged direction.
This could introduce aliases on the predictions and increase the computational
load for fitting the predictor.

We propose a new source distribution that could simultaneously solve the
problems of the preceding ones: the \emph{block averaged sources}.
It consist in dividing the region in blocks of equal size, computing the median
location of the observation points that fall inside each block, and putting one
point source bellow this block averaged coordinate
(fig.~\ref{fig:block-median-sources}).
It creates one source beneath each populated block, solving the problem
of the \emph{grid sources} layout, but also reducing the number of point
sources in comparison with the \emph{source beneath data} layout.

\begin{figure*}
    \includegraphics[width=\linewidth]{figs/block-median-sources-schematics.pdf}
    \caption{
        Block averaged sources. Given a set of observation points (a) we divide
        the survey region on blocks of equal size (b) and compute the median
        location of the observation points that fall inside each block.
        A source point is located beneath each block averaged coordinate (c).
        The number of point sources could be less or equal to the number of
        observation points.
        No source is located beneath unpopulated blocks.
    }
    \label{fig:block-median-sources}
\end{figure*}

The depth of the point sources can be chosen following different criteria.
Deep sources generate low frequencies, while shallow ones generate high
frequencies (cite).
The most simple option is to locate all point sources at the same depth, which
we will call \emph{constant depth} in the future.
If the measurement were taken at significantly different altitudes, the
elevated computation points will be more distant to the sources than the lower
ones.
This may create problems on reproducing high frequencies on the elevated
points.

One possible solution is to locate each source bellow its corresponding
observation point or block averaged point at a constant \emph{relative depth}
from it.
The sources won't be all at the same depth, but they will all be at the same
distance from their corresponding observation point (or block averaged point).
It worth mention that this approach could not be applied to \emph{grid
sources}.

In case our survey presents heavily clustered data points, we may want to
reduce the depth of the sources bellow that region in order to reproduce the
high frequencies measured by these clustered observation points.
We can propose a \emph{variable depth} approach: it locates each point source
according to the \emph{relative depth} approach and then add a term
proportional to the mean distance to its $k$ nearest neighbour sources.
So, the depth of the sources can be computed as follows:

@@

\noindent where \emph{data point height} is the elevation of the corresponding
observation (or block averaged) point to the source, \emph{depth factor} and
\emph{relative depth} are parameters and \emph{mean distance} is the mean
distance to the $k$ nearest source neighbours.

@@

>>>>

How to choose how many sources to use and where to locate them are not
trivial decisions that play an important role on the accuracy of the
predictions and the computational resources needed to fit the sources'
coefficients.

@@

A high number of sources homogeneously distributed on the survey region are
capable of reproducing the observed data. Nevertheless, too many sources will
make the Jacobian matrix too large, what could introduce computational problems
like the inability of storing the matrix on memory or a high computational time
needed to fit the sources' coefficients.
On the other hand, using very few sources will undoubtedly reduce the memory
needs and lower the computational time when fitting the sources' coefficients.
Nevertheless, it could be incapable of making accurate predictions or even
reproducing the measured data if the number of sources is too low.

The choice of the source distribution could also depend on the characteristic
of the survey and how the observation points are distributed on the region.
A ground survey usually consists in observation points located along irregular
paths plus some possible scattered points.
The survey region is usually heterogeneously covered, leaving areas without any
observation.
On the other hand, an airborne survey consists in observation points located
along almost straight flight lines. The measurements are usually taken at
a high frequency, so the observation points along the flight lines are very
close to each other.
This creates an anisotropy on the observation points: there are much more
observation points along the flight lines directions.

@@

The \emph{regular grid} creates an homogeneous distribution of point sources
bellow the surveyed region.
The distance between sources is controlled by the grid spacing parameter.
Keeping the boundaries of the grid constant, reducing the spacing increases the
number of sources.

The \emph{source beneath data} layout locates one source bellow each
observation point, therefore the number of sources will be equal to the number
of data points.

When working with ground surveys, the \emph{regular grid} needs the grid
spacing to be sufficiently small so observed values can be accurately
reproduced by the source distribution.
This will create too many sources on areas where no observation has been
carried out.
In contrast, the \emph{source beneath data} layout is more likely to accurately
fit the observed data with much less sources, reducing the computational load
to fit the equivalent sources' coefficients.

When working with airborne surveys, the \emph{source beneath data} may
generate an anisotropic distribution of sources by locating too many sources
along the flight paths.
It could lead to unwanted aliases on the predicted values.
In contrast, the \emph{regular grid} won't generate aliases by homogeneously
locating sources.

We propose a new source distribution that could simultaneously reduce the
computational load and remove the drawbacks of the preceding ones: the
\emph{block averaged sources}.
It consist in:

\begin{enumerate}
    \item Dividing the region in blocks of equal size. The size of the blocks
        is controlled by a \emph{spacing} parameter.
        (see Fig.~\ref{fig:block-median-sources}(a)-(b)).
    \item Computing the median location of the observation points that fall
        inside each block. Blocks without any observation point are ignored.
    \item Locating one point source bellow each block averaged coordinate.
        (see Fig.~\ref{fig:block-median-sources}(c)).
\end{enumerate}

The number of sources created by this new distribution will be less or equal to
the ones created by the \emph{source beneath data} and the \emph{regular
grid} with the same spacing.
Moreover, the sources are only located beneath populated blocks, so sources
will be concentrated on regions where data has been gathered and no anisotropic
distribution of sources will be generated.

\begin{figure*}
    \includegraphics[width=\linewidth]{figs/block-median-sources-schematics.pdf}
    \caption{
        Block averaged sources. (a) Given a set of observation points (b) we divide
        the survey region on blocks of equal size and compute the median
        location of the observation points that fall inside each block.
        (c) A source point is located beneath each block averaged coordinate.
        The number of point sources could be less or equal to the number of
        observation points.
        No source is located beneath unpopulated blocks.
    }
    \label{fig:block-median-sources}
\end{figure*}

The depth of the point sources can be chosen following different criteria.
Deep sources generate low frequencies, while shallow ones generate high
frequencies (cite).
The simplest option is to locate all sources at the same depth, which we will
call \emph{constant depth} in the future (see Fig.~\ref{fig:depth-types}(a)).
If the measurement were taken at significantly different altitudes, the
elevated computation points will be more distant to the sources than the lower
ones.
This may create problems on reproducing high frequencies on the elevated
points.

One possible solution is to locate each source bellow its corresponding
observation (or block averaged) point at a constant \emph{relative depth}
from it.
The depth of each source can be computed as the height of its corresponding
observation (or block averaged) point minus a depth parameter that assumes the
same value for all sources (see Fig.~\ref{fig:depth-types}(b)).
The sources won't be all at the same depth, but they will all be at the same
distance from their corresponding observation (or block averaged) point.

In case our survey presents heavily clustered data points on some areas, we may
want the sources bellow that region to be shallower in order to reproduce the
high frequencies measured by these clustered observation points.
We can propose a \emph{variable depth} approach: it locates each point source
according to the \emph{relative depth} and then subtracts a term proportional
to the mean distance to its $k$ nearest neighbour sources
(see Fig.~\ref{fig:depth-types}(c)).
So, the depth of the sources can be computed as follows:

@@

\noindent where \emph{data point height} is the elevation of the corresponding
observation (or block averaged) point to the source, \emph{depth factor} and
\emph{relative depth} are parameters and \emph{mean distance} is the mean
distance to the $k$ nearest source neighbours.
This term will be smaller in case of clustered sources, making them shallower
than scattered and distant sources.

@@

Worth mention that the two last approaches (the \emph{relative depth} and the
\emph{variable depth}) could not be applied to \emph{grid sources}.
